---
title: "PML: Project"
output: 
  html_document:
    keep_md: true
---

####Summary####
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset). 

####Load Libraries####
```{r, warning=FALSE, message=FALSE  }
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

setwd("H:/Courses/DataScience/PML/Project/PML_Project")

####Download data####
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "H:/Courses/DataScience/PML/Project/PML_Project/pml-training.csv"
testFile  <- "H:/Courses/DataScience/PML/Project/PML_Project/pml-testing.csv"
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}


####Load data####
```{r}
TrainData = read.csv("pml-training.csv", na.strings = c("NA", ""))
dim(TrainData)
TestData=read.csv("pml-testing.csv", na.strings = c("NA", ""))
TrainData$classe <- as.factor(TrainData$classe)  
```

####Process data####
We need to remove incomplete rows from the dataset.
We also remove the first 7 columns as they are not predictors.
```{r ProcessData, echo = TRUE}
na_test = sapply(TrainData, function(x) {sum(is.na(x))})
tbl = table(na_test)
sum(complete.cases(TrainData))

bad_columns = names(na_test[na_test==19216])
TrainDataClean = TrainData[, !names(TrainData) %in% bad_columns]
TrainData = TrainData[,-c(1:7)]
str(TrainDataClean)
```

Then we apply same transformations on the testdata.
```{r , echo = TRUE}
na_test = sapply(TestData, function(x) {sum(is.na(x))})
tbl = table(na_test)

bad_columns = names(na_test[na_test==20])
TestDataClean = TestData[, !names(TestData) %in% bad_columns]
TestDataClean = TestDataClean[,-c(1:7)]
```

####Partition train data####
We split the training set 70%-30% for cross validation purposes.

```{r PartitionData, echo = TRUE}
set.seed(308)
in_train <- createDataPartition(TrainDataClean$classe, p = 0.7, list = FALSE)
TestSet <- TrainDataClean[-in_train,]
TrainSet <- TrainDataClean[in_train,]

```


####Model data ####
First we use Random Forest package withoutCaret package to model our data.

```{r ModelData, echo = TRUE}
modelRF = randomForest(classe~., data=TrainSet)
modelRF

```


####Cross Validation####

Both models provide good results:
Random Forest:
```{r , echo = TRUE}

mean(predict(modelRF, TestSet) == TestSet$classe) * 100
```

We perform cross validation on Random Forest model 
```{r , echo = TRUE}
caretmodelRF = readRDS("caretmodelRF.RDS")
predictcaretRf <- predict(caretmodelRF, TestSet)
confusionMatrix(TestSet$classe, predictcaretRf)
accuracy <- postResample(predictcaretRf, TestSet$classe)
```

####Predict ####

Using using Random Forest model we predict test data.
```{r PredictData, echo = TRUE}

result <- predict(modelRF, TestDataClean)

result
```
